---
output: 
  html_document:
    css: www/styles_treegrowth.css
    anchor_sections: FALSE
    includes:
        in_header: "header_web.html"
    self_contained: yes
---
## Tree Growth and Mortality plots for Acadia National Park in 2026 {.tabset .tabset-pills}

```{r echo = F, results = 'hide'}
knitr::opts_chunk$set(echo = FALSE, fig.align = 'center', warning = FALSE, message = F, fig.width = 12, fig.height = 8)
options(knitr.kable.NA = '-')
```

```{r, echo = F, results = 'hide'}
library(forestNETN)
library(patchwork)
library(tidyverse)
library(leaflet)

#importCSV("C:/Temp/", zip_name = "NETN_Forest_20260701.zip")
importData(instance = "server", server = "inp2300irmadb01\\ntwk") 
report_year = as.numeric(2026)
parkcode = "ACAD"
rend = TRUE
```

```{r}
subunits <- c("MDI_East", "MDI_West", "Schoodic", "Isle_au_Haut")

top_tree_spp <- joinTreeData(park = parkcode, from = 2006, to = report_year, status = 'live') |> 
  mutate(num_trees = n()) |> 
  summarize(num_species = n(),
            pct_total = (num_species/first(num_trees))*100,
            .by = ScientificName) |> arrange(desc(pct_total)) |> 
  filter(pct_total > 0.1) |> 
  select(species = ScientificName) |> arrange(species)

mdi_e_list <- joinLocEvent(park = parkcode, from = 2006, to = report_year) |> filter(ParkSubUnit == "ACAD_MDI_East") |>   
  select(Plot_Name) |> distinct() 

mdi_w_list <- joinLocEvent(park = parkcode, from = 2006, to = report_year) |> filter(ParkSubUnit == "ACAD_MDI_West") |>   
  select(Plot_Name) |> distinct() 

sch_list <- joinLocEvent(park = parkcode, from = 2006, to = report_year) |> filter(ParkSubUnit == "ACAD_Schoodic") |>   
  select(Plot_Name) |> distinct() 

iah_list <- joinLocEvent(park = parkcode, from = 2006, to = report_year) |> filter(ParkSubUnit == "ACAD_Isle_au_Haut") |>   
  select(Plot_Name) |> distinct() 

# plot_list <- joinLocEvent(park = parkcode, from = 2006, to = report_year) |> select(Plot_Name) |> distinct() 

```


### Background
The plots in this summary show individual trees that were alive the first time they were recorded on a plot, up through the first time a tree was recorded as dead. Each line on the plot indicates a tree and its diameter at breast height (DBH) as it has changed over time. Each tick '|' or 'X' is a sample event. Lines are color coded by growth rate, with darker green indicating faster growth, orange indicating no growth between surveys, and red indicating shrinkage between surveys. See <a href="https://katemmiller.shinyapps.io/NETN_FHM_photos">NETN photo viewer</a> for plot locations.  

<h3>Forest Plot Map</h3>
```{r}
forest_plots <- joinLocEvent(park = "ACAD", output = 'verbose') |> 
  select(Plot_Name, PlotCode, ParkSubUnit, PhysioClass = PhysiographySummary, Directions, PlotNotes, Lat, Long) |> 
  distinct()
```

```{r eval = F}
# Run once to get all the plots. 
tree_plots <- purrr::map(forest_plots$Plot_Name, function(p){
   plotTreeGrowth(plotName = p)
   ggsave(paste0("./output/2026/tree_growth_plots/", p, ".png"), width = 6, height = 4) 
})
```

```{r}
# map tiles urls
ESRIimagery <- "http://services.arcgisonline.com/arcgis/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}"
ESRItopo <- "http://services.arcgisonline.com/arcgis/rest/services/World_Topo_Map/MapServer/tile/{z}/{y}/{x}"
ESRINatGeo <- "http://services.arcgisonline.com/arcgis/rest/services/NatGeo_World_Map/MapServer/tile/{z}/{y}/{x}"

forest_map <- 
leaflet(height = 800, width = "90%") |> 
    # set default view of park
    setView(lng = -68.312,
            lat = 44.25,
            zoom = 10.25) |> 
    setMaxBounds(lng1 = -69,
                 lng2 = -67.5,
                 lat1 = 44.9,
                 lat2 = 43.5) |> 
    # add map tiles
    addTiles(group = "Topo", urlTemplate = ESRItopo, options = tileOptions(minZoom = 8)) |>
    addTiles(group = "Imagery", urlTemplate = ESRIimagery, options = tileOptions(minZoom = 8)) |>
    addTiles(group = "NatGeo", urlTemplate = ESRINatGeo, options = tileOptions(minZoom = 8)) |>
    addLayersControl(
      baseGroups = c("Topo", "Imagery", "NatGeo"),
      options = layersControlOptions(collapsed = T)) |> 
    # add button to control map tiles
    addLayersControl(baseGroups = c("Topo", "Imagery", "NatGeo"),
                     options = layersControlOptions(collapsed = T)) |> 
    # add site markers 
    clearPopups() |> 
    clearControls() |>
    addCircleMarkers(data = forest_plots,
                     radius = 4,
                     lng = forest_plots$Long,
                     lat = forest_plots$Lat,
                     layerId = forest_plots$PlotCode,
                     label = forest_plots$PlotCode,
                     labelOptions = labelOptions(noHide = T, textOnly = TRUE, direction = 'bottom', textsize = "10px"),
                     fillColor = '#66bd63',
                     fillOpacity = 0.75,
                     weight = 1.5,
                     color = "DimGrey",
                     popup = paste0('<b>', forest_plots$Plot_Name, "</b><br>",
                                    '<b>Physiographic: </b>', forest_plots$PhysioClass, "<br>",
                                    '<b>Plot Note: </b>', forest_plots$PlotNotes, "<br>",
                                    '<b>Directions: </b>', forest_plots$Directions, "<br>",
                                    paste0('<img src = "./output/2026/tree_growth_plots/', forest_plots$Plot_Name, '.png">', 
                                           ' alt = ', '"Tree growth plot for plot ', forest_plots$Plot_Name), '"<br>')
                     )

forest_map

```

### Plots with >5% mortality {.tabset}
#### MDI-E {.tabset}
<details open><summary class = 'drop'>Panel 1</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 1, title = F, elev_mort = T)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 2</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 2, title = F, elev_mort = T)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 3</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 3, title = F, elev_mort = T)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 4</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 4, title = F, elev_mort = T)
```
</details>
<br>

#### MDI-W {.tabset}
<details open><summary class = 'drop'>Panel 1</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 1, title = F, elev_mort = T)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 2</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 2, title = F, elev_mort = T)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 3</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 3, title = F, elev_mort = T)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 4</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 4, title = F, elev_mort = T)
```
</details>
<br>

#### Schoodic {.tabset}
```{r}
plotTreeGrowth(park = "ACAD", subunit = "Schoodic", title = F, elev_mort = T)
```

#### IAH {.tabset}
```{r}
plotTreeGrowth(park = "ACAD", subunit = "Isle_au_Haut", title = F, elev_mort = T)
```


### Species-level plots {.tabset}
```{r, results = 'asis', eval = rend}
for(i in seq_along(top_tree_spp$species)){
  spp <- top_tree_spp$species[[i]]
  cat("#### ", spp, "\n\n")
  sub <- "MDI_East"
  cat("<details open><summary class = 'drop'>MDI-E</summary>", "\n\n")
  cat("\n\n")
  tryCatch(print(plotTreeGrowth(park = "ACAD", species = spp, subunit = sub)),
         error = function(e){cat("No trees for given species in specified unit.")})
  cat("\n\n")
  cat("</details>", "\n\n")
  
  sub <- "MDI_West"
  cat("<details open><summary class = 'drop'>MDI-W</summary>", "\n\n")
  cat("\n\n")
  tryCatch(print(plotTreeGrowth(park = "ACAD", species = spp, subunit = sub)),
           error = function(e){cat("No trees for given species in specified unit.")})
  cat("\n\n")
  cat("</details>", "\n\n")

  sub <- "Schoodic"
  cat("<details open><summary class = 'drop'>Schoodic</summary>", "\n\n")
  cat("\n\n")
  tryCatch(print(plotTreeGrowth(park = "ACAD", species = spp, subunit = sub)),
           error = function(e){cat("No trees for given species in specified unit.")})
  cat("\n\n")
  cat("</details>", "\n\n")

  sub <- "Isle_au_Haut"
  cat("<details open><summary class = 'drop'>IAH</summary>", "\n\n")
  cat("\n\n")
  tryCatch(print(plotTreeGrowth(park = "ACAD", species = spp, subunit = sub)),
           error = function(e){cat("No trees for given species in specified unit.")})
  cat("\n\n")
  cat("</details>", "\n\n")

  cat("\n\n")

}
  
```

### All plots and trees {.tabset}
#### MDI-E {.tabset}
<details open><summary class = 'drop'>Panel 1</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 1, title = F)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 2</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 2, title = F)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 3</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 3, title = F)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 4</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_East", panels = 4, title = F)
```
</details>
<br>

#### MDI-W {.tabset}
<details open><summary class = 'drop'>Panel 1</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 1, title = F)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 2</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 2, title = F)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 3</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 3, title = F)
```
</details>
<br>
<details open><summary class = 'drop'>Panel 4</summary>
```{r, fig.height = 12}
plotTreeGrowth(park = "ACAD", subunit = "MDI_West", panels = 4, title = F)
```
</details>
<br>

#### Schoodic
```{r}
plotTreeGrowth(park = "ACAD", subunit = "Schoodic", title = F)
```

#### IAH
```{r}
plotTreeGrowth(park = "ACAD", subunit = "Isle_au_Haut", title = F)
```


